---
title: "MP2"
author: "Cindy Sun, Nora Zhou, Sarah Susnea"
format: pdf
---

```{r setup, include=FALSE}
# Set default behavior for all code chunks here:
knitr::opts_chunk$set(
  echo = TRUE, 
  message = FALSE, 
  warning = FALSE, 
  fig.width = 16/2, 
  fig.height = 9/2
)

library(dplyr)
library(tidyr)
library(ggplot2)
library(knitr)
library(readr)

data_raw <- read.csv("data_raw.csv")
```

# Introduction

What motivates college students to enroll in a particular class? This study seeks to investigate the correlation between the timing of class offerings and college students' inclination to register for the class. The hypothesis is that college students exhibit preferences or aversions towards certain class schedule times. Specifically, it is expected that students generally favor morning or afternoon classes over evening ones.

# Method 

The study collected data specifically for the SDS290 class during the Spring of 2024, with a total of 37 college students participating. Two incomplete responses were removed during the data cleaning process, resulting in a dataset of 35 observations. The experiemntal unit of this study is participants. Each participant was presented with a scenario regarding course registration and then asked to rate their interest in enrolling in the class on a scale of 1 to 5, which is the response variable of this study. The scenario provided was as follows:


*You've come across a fascinating course that you'd like to take next semester, but unfortunately, Smith does not offer this particular class. Luckily, you discovered that Amherst College provides this course, scheduled every Monday and Wednesday from* ___ *to* ___*. On a scale of 1 to 5, how inclined are you to enroll in this course? (1 = not interested at all, 5 = Yes! I will definitely take it)*


The blanks in the scenario represent class schedules, which is the factor of interest of this study and are divided into three levels. Each participant was randomly assigned to one of these three levels: 

- scheduled every Monday and Wednesday from 10:10 am to 11:30 am (ConditionMorning).

- scheduled every Monday and Wednesday from 2:10 pm to 3:30 pm (ConditionAfternoon).

- scheduled every Monday and Wednesday from 4:50 pm to 6:10 pm (ConditionEvening).

# Results
```{r, echo=FALSE}
# data cleaning
data_raw <- data_raw[-1:-2, ]

data_clean <- data_raw |>
  filter(Finished != "FALSE")|>
  filter(Consent != "No") |> 
  select(-(1:(ncol(data_raw) - 4)))|>
  pivot_longer(cols = c(QuestionMorning, QuestionAfternoon, QuestionEvening), 
               names_to = "condition", values_to = "score") |>
  filter(!is.na(score))

data_clean$score <- as.numeric(data_clean$score)

data_clean <- data_clean |>
  filter(!is.na(score))

names <- c("car", "condition", "score")
colnames(data_clean) <- names

# write the cleaned data into csv for submission
write_csv(data_clean, file = "clean_data.csv")
```

```{r}
ggplot(data_clean, aes(x = condition, y = score, 
                    group = car, 
                    color = car)) +
  geom_point() +
  geom_smooth(method = "lm", se = 0)
```

```{r}

```



## Conditions Check
```{r, echo=FALSE}
# fit the model
mod <- lm(score ~ condition*car, data = data_clean)
```

- C. Constant effects: 
Assumed each participant is affected exactly the same in each condition group.

- A. Additive effects:
Assumed effects of condition, car and interaction of condition and car are addictive.

- S. Same standard deviations:
```{r, echo=FALSE}
# descriptive data
data_clean |>
    group_by(condition) |>
    summarize(sd = sd(score)) |>
    summarize(max(sd)/min(sd))

data_clean |>
    group_by(car) |>
    summarize(sd = sd(score)) |>
    summarize(max(sd)/min(sd))

data_clean |>
    group_by(condition, car) |>
    summarize(sd = sd(score)) |>
    ungroup() |>
    summarize(max(sd)/min(sd))

plot(mod, which = 1)
```

The condition is met because the largest standard deviation is not yet 2 times of the smallest standard deviation (1.21/1.13 = 1.070796).

- I. Independent residuals:
Assumed the residuals are independent due to independent participants. 

- N. Normally distributed residuals:
```{r, echo=FALSE, fig.height=3, fig.width=5}
plot(mod, which = 2)
```
Residual distribution is roughly normal, except for those deviate at both ends. However, since real-world data wouldn't be perfect and there are only 39 complete responses in this study, the condition is assumed satisfied. 

- Z. Zero mean residuals:
```{r, echo=FALSE, fig.height=3, fig.width=5}
qplot(mod$residuals, bins = 5)
```
According to the histogram, the distribution is normal with the mean of the residuals centered at zero. This condition is satisfied.

## ANOVA table and boxplot
After checking all the conditions, the model has been fitted and an ANOVA test has been used. The ANOVA table is provided below:
```{r, echo=FALSE}
anova(mod)
```

A boxplot showing the variance of means of the three condition groups is shown below:
```{r, echo=FALSE,fig.height=3, fig.width=5}
ggplot(data = data_clean, 
       aes(x =condition, y = score, fill = car)) + 
       geom_boxplot()
```

According to both the ANOVA result and the boxplot, there is a statistically significant effect of course schedules on college students' willingness to enroll in a class (p=0.01, which is smaller than 0.05). The hypothesis of college students exhibit preferences or aversions towards certain class schedule times is favored. Significantly, the boxplot demonstrates a greater average willingness score for class offerings in the afternoon compared to both evening and morning sessions.

## Fisher's LSD 
```{r, echo=FALSE}
MSE = 1.3881
df_E = 32
t <- qt(.975, df_E)

# LSD for ConditionMorning and ConditionAfternoon
LSD1 = t*sqrt(MSE)*sqrt(1/11+1/13)

# LSD for ConditionAfternoon and ConditionEvening
LSD2 = t*sqrt(MSE)*sqrt(1/13+1/11)

# LSD for ConditionMorning and ConditionEvening
LSD3 = t*sqrt(MSE)*sqrt(1/11+1/11)

# Mean difference between ConditionMorning and ConditionAfternoon
mean1 <- 3.45-2
# Mean difference between ConditionAfternoon and ConditionEvening
mean2 <- 3.45-2.15
# Mean difference between ConditionMorning and ConditionEvening
mean3 <- 2.15-2
```

Given the mean squared error (1.3881), the degree of freedom of error (32), and the number of observations in each condition groups, LSD scores can be computed to examine the differences between condition groups. Since this is an unbalanced design with ConditionMorning having 11 observations, ConditionAfternoon having 11 observations, and ConditionEvening having 13 observations, 3 LSD scores and 3 differences in means are computed and compared:

```{r, echo=FALSE}
# create a data frame to store LSD and mean differences
df <- data.frame(
  Comparison = c("ConditionMorning and ConditionAfternoon", 
                 "ConditionAfternoon and ConditionEvening", 
                 "ConditionMorning and ConditionEvening"),
  LSD = c(LSD1, LSD2, LSD3),
  MeanDifference = c(mean1, mean2, mean3),
  Significant = c("Yes", "Yes", "No")
)

# Create a table using kable
table <- kable(df)
table
```

According to the table reporting the comparison between the LSD and the mean difference score for each comparison, two comparisons were found to have significant different means (MeanDifference is higher than LSD score). Only ConditionMorning and ConditionEvening have insignificant difference in mean (MeanDifference is lower than LSD score). 


## Confidence Interval
Confidence intervals are computed for all three comparison groups in the table below:
```{r, echo=FALSE}
# calculate lower and upper bounds for each comparison
LL1 <- mean1 - LSD1 #lower limit
UL1 <- mean1 + LSD1 #upper limit

LL2 <- mean2 - LSD2 #lower limit
UL2 <- mean2 + LSD2 #upper limit

LL3 <- mean3 - LSD3 #lower limit
UL3 <- mean3 + LSD3 #upper limit

# create a data frame to store confidence intervals
df2 <- data.frame(
  Comparison = c("ConditionMorning and ConditionAfternoon", 
                 "ConditionAfternoon and ConditionEvening", 
                 "ConditionMorning and ConditionEvening"),
  LowerLimit = c(LL1, LL2, LL3),
  UpperLimit = c(UL1, UL2, UL3),
  Significant = c("Yes", "Yes", "No")
)

# Create a table using kable
table <- kable(df2)
table
```

According to the table reporting the lower and upper bounds of confidence intervals for each comparison groups, two comparisons were found to have significant different means (confidence interval does not contain zero). Only ConditionMorning and ConditionEvening have insignificant difference in mean (confidence interval contains zero). This result aligns with the previous result from the Fisher's LSD test.

## Effect Size (D)
After two comparison groups have been tested to have significant difference in means, the effect size (D) is measured to examine how large the condition effects are. A table of effect size is reported below:

```{r, echo=FALSE}
# calculate the effect size scores
D1 = mean1/sqrt(MSE)
D2 = mean2/sqrt(MSE)

# create a data frame to store effect size scores
df3 <- data.frame(
  Comparison = c("ConditionMorning and ConditionAfternoon", 
                 "ConditionAfternoon and ConditionEvening"),
  D = c(D1, D2)
)

# Create a table using kable
table <- kable(df3)
table
```

According to the table, both comparison groups exhibit substantial effect sizes (D > 1). Specifically, the disparity in willingness between students attending morning and afternoon classes is 1.23 times the magnitude of typical within-group deviations in willingness. Similarly, the variance in willingness between students attending afternoon and evening classes is 1.10 times the magnitude of typical within-group deviations in willingness.

## R-squared
```{r, echo=FALSE}
R2 = 14.323/(14.323+44.420)
```
Given the sum of squares due to regression (SSR) and the sum of squares total (SST) scores from the ANOVA table, the R-squared value is computed to be 0.2438248. This means that about 24% of the variation in college students' willingness of taking the course is explained by when the class is offered.



# Conclusion
This research aims to investigate the correlation between the timing of class offerings and college students' inclination to enroll in the class. The hypothesis is that students will display preferences or aversions towards specific class schedule times, particularly favoring morning or afternoon classes over evening ones. The ANOVA test results reveal a statistically significant impact of class timing on students' inclination to register, which favors the hypothesis. Furthermore, the evaluation of Fisher's LSD scores and confidence intervals indicate significant mean differences between two out of three comparison groups (ConditionMorning and ConditionAfternoon; ConditionAfternoon and ConditionEvening). This suggests a preference among college students for courses offered in the afternoon over those offered in the morning or evening, with both groups exhibiting large effect sizes. The R-squared value of 0.2438248 signifies that approximately 24% of the variation in students' willingness to enroll in the course can be attributed to the timing of class offerings, indicating a moderate explanatory power.

Despite the significant findings, several limitations of the study should be acknowledged. Firstly, the sample size employed in the research is relatively small and limited to the SDS 290 class, which may affect the generalizability of the results to a broader population of college students. Secondly, while the timing of class offerings explains 24% of the variation in students' enrollment preferences, it is important to recognize that other factors or confounding variables may play a role in influencing these preferences to a greater extent. Therefore, future research should aim to address these limitations by employing larger sample sizes, adopting a balanced experimental design, and exploring additional variables that could contribute to a more comprehensive understanding of college students' class enrollment behaviors.











