---
title: "MP2"
author: "Cindy Sun, Sarah Susnea, Nora Zhou"
format: pdf
---

```{r setup, include=FALSE}
# Set default behavior for all code chunks here:
knitr::opts_chunk$set(
  echo = TRUE, 
  message = FALSE, 
  warning = FALSE, 
  fig.width = 16/2, 
  fig.height = 9/2
)

library(dplyr)
library(tidyr)
library(ggplot2)
library(knitr)
library(readr)

data_raw <- read.csv("data_raw.csv")
```

# Introduction

What drives college students to choose specific classes? This research aims to explore how the timing of class schedules and car ownership influence college students' likelihood of enrolling in a particular course. We hypothesized that there is a main effect of the timing of classes on students' registration preferences, with a preference for afternoon sessions over morning and evening ones. Additionally, we hypothesized that car ownership plays a role, with students who own cars being more inclined to register for the class. We also hypothesized that there will be a significant interaction effect of class schedules and car ownership on preferences of class registration.

# Method 

The study collected data specifically for the SDS290 class during the Spring of 2024, with a total of 38 college students participating. One incomplete response was removed during the data cleaning process, resulting in a dataset of 38 observations. The experimental unit of this study is participant. Each participant was presented with a scenario regarding course registration and then asked to rate their interest in enrolling in the class on a scale of 1 to 5, which is the response variable of this study. The scenario provided was as follows:


*You've come across a fascinating course that you'd like to take next semester, but unfortunately, Smith does not offer this particular class. Luckily, you discovered that Amherst College provides this course, scheduled every Monday and Wednesday from* ___ *to* ___*. On a scale of 1 to 5, how inclined are you to enroll in this course? (1 = not interested at all, 5 = Yes! I will definitely take it)*


The blanks in the scenario represent class schedules, which is the factor of interest of this study and are divided into three levels. Each participant was randomly assigned to one of these three levels: 

- scheduled every Monday and Wednesday from 10:10 am to 11:30 am (ConditionMorning).

- scheduled every Monday and Wednesday from 2:10 pm to 3:30 pm (ConditionAfternoon).

- scheduled every Monday and Wednesday from 4:50 pm to 6:10 pm (ConditionEvening).

In addition to the scenario question, all participants were asked if they own a car on campus. They could either answer "Yes" or "No."

# Results
```{r, echo=FALSE}
# data cleaning
data_raw <- data_raw[-1:-2, ]

data_clean <- data_raw |>
  filter(Finished != "FALSE")|>
  filter(Consent != "No") |> 
  select(-(1:(ncol(data_raw) - 4)))|>
  pivot_longer(cols = c(QuestionMorning, QuestionAfternoon, QuestionEvening), 
               names_to = "condition", values_to = "score") |>
  filter(!is.na(score))

data_clean$score <- as.numeric(data_clean$score)

data_clean <- data_clean |>
  filter(!is.na(score))

names <- c("car", "condition", "score")
colnames(data_clean) <- names

# write the cleaned data into csv for submission
write_csv(data_clean, file = "clean_data.csv")
```

## Conditions Check
```{r, echo=FALSE}
# fit the model
mod <- lm(score ~ condition*car, data = data_clean)
```

- C. Constant effects: 
Assumed each participant is affected exactly the same in each condition group.

- A. Additive effects:
Assumed effects of condition, car and interaction of condition and car are additive.

- S. Same standard deviations:


The max sd divided by the min sd, grouped by class schedules, is calculated below:
```{r, echo=FALSE}
# descriptive data
data_clean |>
    group_by(condition) |>
    summarize(sd = sd(score)) |>
    summarize(max(sd)/min(sd))
```
According to the result above, the condition is met for the main effect of class schedules on registration preferences.


The max sd divided by the min sd, grouped by car ownership, is calculated below:
```{r, echo=FALSE}
data_clean |>
    group_by(car) |>
    summarize(sd = sd(score)) |>
    summarize(max(sd)/min(sd))
```
According to the result above, the condition is met for the main effect of car ownership on registration preferences.


The max sd divided by the min sd, grouped by class schedules & car ownership, is calculated below:
```{r, echo=FALSE}
data_clean |>
    group_by(condition, car) |>
    summarize(sd = sd(score)) |>
    ungroup() |>
    summarize(max(sd)/min(sd))
```

According to the result above, the condition is violated for the interaction effect of class schedules & car ownership on registration preferences. The largest SD is slightly more than two times (2.12) as big as the smallest for the interaction condition. 


We applied the log transformation on our outcome variable, and reevaluated the same standard deviation condition, the max sd divided by the min sd of the interaction effect is given below:
```{r, echo=FALSE}
data_clean |>
    group_by(condition, car) |>
    summarize(sd = sd(log(score))) |>
    ungroup() |>
    summarize(max(sd)/min(sd))
```
We see that after log transformation, the difference between standard deviations are even larger (2.81>2.12). Therefore, we chose to not transform our outcome variable in model fitting. 


Based on the calculation, our condition of same standard deviation is slightly violated. However, when we plot the residual plot (see below), we see that the red line falls along y = 0. Therefore, we still assume that this condition is met. 
```{r, echo=FALSE}
plot(mod, which = 1)
```


- I. Independent residuals:
Assumed the residuals are independent due to independent participants. 

- N. Normally distributed residuals:
```{r, echo=FALSE, fig.height=3, fig.width=5}
plot(mod, which = 2)
```
Residual distribution is roughly normal, except for those deviate at both ends. However, since real-world data wouldn't be perfect and there are only 38 complete responses in this study, the condition is assumed satisfied. 

- Z. Zero mean residuals:
```{r, echo=FALSE, fig.height=3, fig.width=5}
qplot(mod$residuals, bins = 5)
```
According to the histogram, the distribution is normal with the mean of the residuals centered at zero. This condition is satisfied.

## Plots and ANOVA table

A boxplot showing the variances and mediums of the three condition and two car groups is shown below:
```{r, echo=FALSE,fig.height=3, fig.width=5}
ggplot(data = data_clean, 
       aes(x =condition, y = score, fill = car)) + 
       geom_boxplot()
```

An plot showing the main effects of condition and car ownership and the interaction between them on registration preferences are shown below:

```{r,echo = F,fig.height=3, fig.width=5}
ggplot(data_clean, aes(x = condition, y = score, 
                    group = car, 
                    color = car)) +
  geom_point(alpha = 0.2) +
  geom_smooth(method = "lm", se = 0)
```

After checking all the conditions, the model has been fitted and an ANOVA test has been used. The ANOVA table is provided below:
```{r, echo=FALSE}
anova(mod)
```

Surprisingly, based on the boxplot, people who have cars on campus are less likely to register for the class than their counterparts. However, according to the ANOVA table, there is no statistically significant main effect of car ownership on class registration preferences F(1, 32) = 2.1518, p = 0.1522.

In addition, there is no statistically significant main effect of condition on college students' willingness to enroll in a class, F(2, 32) = 0.1630, p = 0.8503. Neither does interaction between condition and car, F(2, 32) = 0.0443, p = 0.9567. Therefore, we reject all of our hypotheses.


## R-squared
```{r, echo=FALSE}
R2 <- (0.428 + 2.824 + 0.116)/(0.428 + 2.824 + 0.116 + 42.000)
```
Given the sum of squares from the ANOVA table, the R-squared value is computed to be 0.074. This means that only 7.4% of the variation in college students' willingness of taking the course is explained by the model.


# Conclusion
This research aims to investigate the correlation between the timing of class offerings and college students' inclination to enroll in the class. The hypothesis is that students will display preferences or aversions towards specific class schedule times, particularly favoring morning or afternoon classes over evening ones. The ANOVA test results reveal a statistically significant impact of class timing on students' inclination to register, which favors the hypothesis. Furthermore, the evaluation of Fisher's LSD scores and confidence intervals indicate significant mean differences between two out of three comparison groups (ConditionMorning and ConditionAfternoon; ConditionAfternoon and ConditionEvening). This suggests a preference among college students for courses offered in the afternoon over those offered in the morning or evening, with both groups exhibiting large effect sizes. The R-squared value of 0.2438248 signifies that approximately 24% of the variation in students' willingness to enroll in the course can be attributed to the timing of class offerings, indicating a moderate explanatory power.

Despite the significant findings, several limitations of the study should be acknowledged. Firstly, the sample size employed in the research is relatively small and limited to the SDS 290 class, which may affect the generalizability of the results to a broader population of college students. Secondly, while the timing of class offerings explains 24% of the variation in students' enrollment preferences, it is important to recognize that other factors or confounding variables may play a role in influencing these preferences to a greater extent. Therefore, future research should aim to address these limitations by employing larger sample sizes, adopting a balanced experimental design, and exploring additional variables that could contribute to a more comprehensive understanding of college students' class enrollment behaviors.